
In the aftermath of the 2007-2008 global financial crisis (GFC), the use of central clearing counterparties (CCPs) became commonplace for various types of derivative contracts. Today, CCPs have expanded to clearing nearly all kinds of financial instruments, including Cash Equities and Repurchase Agreements. Large US G-SIBs were often referred to as "Too Big to Fail" in the wake of the GFC, a label which arguably now fits these CCPs better. The key role of a CCP is the reduction of counterparty credit risk through novation - they step between each trade and act as the buyer to every seller and the seller to every buyer. With this, the CCP concentrates the counterparty credit risk upon itself, and to protect against this it collects collateral from investors. Risk in this new system is therefore transformed, but it does not disappear. 

The concentration of risk onto the CCP brings concerns of a possibly catastrophic CCP failure, a concern not without precedence. \cite{bignon_vuillemey_2020} list three such failures, each occurring between 1974 and 1984. Recently, a large member default at Swedish Nasdaq Clearing AB in September 2018 led to the exhaustion of 107 million out of 166 million euros in the CCP's default fund. Defaults of a CCP's clearing members (CM) are handled in line with a 'default waterfall'. The first layer of the waterfall is initial margin (IM) and variation margin (VM), the first insuring possible shortfalls while liquidating defaulted position during the margin period of risk, and the second to adjust for mark-to-market changes on the position. The second layer, to be used if the first proves insufficient, is the default fund (DF) contribution of the defaulted member. Third comes a part of the CCP's own contribution to the DF, referred to as CCP skin in the game (SITG). The fourth layer is the DF contribution of non-defaulting members, and if this proves inadequate the CCP enters a state of recovery. 

In addition to the risk of a CCP failure, the collection of this collateral also introduces liquidity risks. These risks are often procyclical -- margins increase due to changes in price or increases in volatility, and the liquidity needs created by higher margins may drive firms to acquire offsetting positions or conduct a fire sale of assets, which lead to price impact driving further margin calls \citep{blanck_2025}. In the early days of the COVID-19 crisis, for reference, IM increased by a third of pre-crisis levels at the four largest UK and EU CCPs due to increased market volatility \citep{esrb_2020_covid}. 

The shift towards central clearing is accompanied by countless studies evaluating the benefits CCPs bring, the liquidity risks that arise from them, the adequacy of margining models and event studies of stress episodes. Studies on the working of the client clearing market, however, remain scarce \citep{menkveld_vuillemey_2021}. Clearing membership at a CCP comes with strict requirements. Members are, therefore, mostly larger banks (BCMs) and specialised non-bank clearing firms (NBCMs), which further offer client clearing services. Other market participants, such as hedge funds, pension funds, small banks and proprietary trading shops use these services. The CM then guarantees the position of its client at the CCP.\\

\noindent This paper proposes an Agent-Based Model (ABM) to explore the role and effect of client clearing on CCP systemic risk, a first of its kind. An ABM offers the ability to recreate the complex and reactive behaviours of market participants and central clearing, and can do so without the need for a proprietary clearing dataset. The ABM is built upon a financial market simulation calibrated to reproduce the empirical stylised facts of returns, linked to a tiered CCP environment that novates these trades. The traders in this financial market simulation consist of BCMs and clients, and the clearing layer of these same BCMs and client traders, and additional NBCMs.

The model in itself serves as the major contribution of this thesis. There is currently no ABM that includes client clearing, and models on or including client clearing are scarce. The model is then used to study margin procyclicality under stressed environments, looking at the allocation of losses under volatility reactive margining, as opposed to flat and constant margining. For this the model shows that margin level relocates default risk between clients and CMs, a low constant margin yields many client defaults that are small in size and absorbed by the CMs. Under a high constant margin, clients default less often but the defaults are larger, causing the default of the CM - mostly NBCMs -  that needs to close out this large position into a falling market. Finally, an adapted version of the model is presented with synthetic stochastic stress paths, allowing for the model's use across various applications and theoretical questions. The calibration of the model's parameters, jointly with the parameters of the synthetic paths is also a novel application of the surrogate modelling technique employed by \cite{gao_2022_xgb}.

The rest of the paper is organised as follows. Chapter~\ref{ch:related} reviews the related literature on agent-based market microstructure, central clearing and CCP systemic risk, and agent-based model calibration. Chapter~\ref{ch:model} sets out the structure of the model: the two-tier architecture, the agent types, the call-auction limit order book, the fundamental value process, and the central-clearing tier with its margin, default-fund, and waterfall mechanics. Chapter~\ref{ch:calibration} presents the empirical data and the calm and stressed regimes, the surrogate-assisted calibration methodology, and the validation of the calibrated market layer against the empirical stylised facts. Chapter~\ref{ch:results} reports the main results, the output of the standard model, its application in studying procyclicality, and the version of the ABM with synthetic paths. Finally, Chapter~\ref{ch:conclusion} concludes, and offers limitations and outlines directions for future work.