% Sections only — \chapter{Related Work}\label{ch:related} is in main.tex.

\section{Clearing and Central Counterparties}\label{sec:rw-clearing}

In their article, "The Economics of Central Clearing", \cite{menkveld_vuillemey_2021} offer a systematic review of the theories and empirical studies on Central Clearing. They state that CCPs create value only under the condition that markets are not perfect. A first economic rationale for CCPs is that they bring efficiency gains in the posting of collateral through higher trustworthiness in its segregation and the benefits of multilateral netting \citep{duffie_zhu_2011}. A second rationale is that CCPs serve as insurance against counterparty defaults through loss mutualisation and pooled resources \citep{biais_heider_hoerova_2012}. A third and last rationale for CCPs is their role in mitigating fire sales - when a member defaults, their positions are auctioned amongst surviving members at prices generally higher than what an open market liquidation would bring \citep{vuillemey_2023}. \cite{menkveld_vuillemey_2021} also highlight research questions that remain open for further work -- including the workings of the client clearing market. 

\citep{borovkova_2013} use a network simulation approach to study the effect of central clearing on the stability of the financial system. They find that while a CCP stabilises homogenous networks, tiered and core-periphery networks that more closely resemble reality, the benefit shrinks or reverses. Furthermore, large clearing members benefit from the risk sharing while smaller members bear increased contagion risk. 

\subsection{Client Clearing}

\cite{ofr_2026} study client clearing and its implications for market structure and stability by looking at Credit Default Swaps. They found that by January 2025, clients accounted for 73\% of margin posted at the largest US and EU CCPs. They find that the modal client has only a single clearing connection, and that client clearing improves dealers' (Clearing Members') netting benefits which also creates pricing benefits for clients. In periods of stress, these dealers may also stop the provision of clearing services - cutting off their clients' access to CCPs. \cite{galbiati_soramaki_2013} build a simplified and stylised network model of tiered CCP-clearer-client relationships. They find that tiering reduces CCP total exposure but concentrates it onto fewer, larger GCMs. These extreme exposures to the CCP become larger and less predictable, despite a reduction in the average large exposure. The effect of concentration onto larger GCMs leads to reduced systemic risk for the CCP - large GCMs can internalise more risk due to netting benefits from offsetting client positions. 

(Rough Draft awaiting results:)
This leads us to the first question explored through the Client Clearing Agent-Based Model -- the comparison of the positive affect of client clearing as a loss absorbing layer against the negative effect due to dependence and concentration towards CCP systemic risk.

\subsection{Margins and Pro-Cyclicality}

Through the advent of central clearing, risk does not disappear; counterparty risk is rather transformed into liquidity risk \citep{cont_2017}. Accounting losses in a bilateral system with Mark-to-Market changes turn into realised margin calls drawing on liquidity in central clearing. As risk-basked margins rise with volatility, these margin calls are pro-cyclical \citep{brunnermeier_pedersen_2009} -- margins peak when funding is low, leading to a spiral which causes forced sales that further depress prices and lead to even more margin calls. \cite{glasserman_wu_2018} find that volatility spikes lead to destabilising margin calls in times of market stress. \cite{esrb_2020} report that at the start of the COVID-19 crisis, IM posted at the four largest EU/UK CCPs rose by a third of pre-crisis levels, daily VM calls at the same time quintupled on euro area investment funds' derivative exposures, and increases at this time stemmed mostly from client portfolios and to a lesser extent from house portfolios. Prompted by the large March 2020 margin calls, \cite{bcbs_cpmi_iosco_2025} set out to review CCP margining practises and created a policy proposal calling for greater transparency in margining practises by CCPs onto CMs and further onto clients. 

An alternative to margins that react to recent market volatility are "through the cycle" margin models that ask for higher margins in quiet times \citep{glasserman_wu_2018}. These additional requirements, also called anti-procyclicality tools, minimise destabilising increases in IM requirements during market stress \citep{blanck_2025}. (Rough Draft awaiting results:) The second question investigated through the model is the effect of pro-cyclical margin calls on systemic risk, and if these risks can be mitigated through more stable margin requirements and anti-procyclicality tools. 

\section{Agent-Based Model for Client Clearing}

Agent-Based modelling is a novel methodology to study complex, adaptive and emergent behaviours arising from the interactions of heterogenous agents \citep{axtell_farmer_2025}. The analysis of these behaviours is particularly useful when looking at stressed market environments, and these models have found increasing popularity in recent years in simulating financial markets. \cite{deloitte_ccp} develop an agent-based CCP risk model to analyse the effects of proposed recommendation to CCP risk practices laid out in a white paper by major global CMs \citep{ccp_path_forward_2020} , such as the adequacy of Cover-2 for CCP default funds. They incorporate a simulated market environment with an exchange and Clearing Members, complete with a limit order book where CMs submit orders and exhibit behaviours found in fundamental, momentum and noise traders. The CCP then collects margins based on these trades, which each CM needs to answer, alongside a default management framework to simulate realistic CCP behaviour. The model in this paper is in many ways an extension of this model, adopting a similar architecture and extending it with a client tier.\footnote{The publicly available descriptions of this model are the
practitioner white paper \citep{deloitte_ccp} and Simudyne's online model
documentation; the detailed model documentation on which
Chapter~\ref{ch:model} draws was provided to the author  by Simudyne and is cited as unpublished technical documentation \citep{simudyne_odd}.}

\subsection{Financial Market Simulations}

The first stage in the model is the design of a simulated market environment. The market simulations attempt to reproduce the stylised facts of returns we see in the market -- heavy tails, near zero autocorrelations and volatility clustering. One approach behind this is to look at price dynamics through linear price impact caused by the aggregate and excess demands of different trader agents. An early paper on this is \cite{chiarella_1992}, which populate the market with fundamentalists trading on a fundamental value and chartists trading on momentum and trend. An extension of this model is proposed by \cite{majewski_2018}, who add noise traders and allow for a long term drift in fundamental asset value and create an Agent-Based Model. In their model, the parameters carried by each trader are jointly estimated alongside the Fundamental Value using a Kalman Filter and an Expectation Maximisation algorithm. \cite{gao_2023} further extend this model by replacing the noise trader with a volatility trader, whose demand level adapts to a volatility signal based on a mean reverting Heston model. The introduction of this element allows for the replication of volatility clustering, which previous models could not accurately replicate. Trader parameters in their model are calibrated by performing a grid search and minimising a loss function that represents distance from stylised facts. 

Another approach replaces the aggregate price impact mechanism with an explicit limit order book (LOB) such that prices emerge from the matching of submitted orders. \cite{chiarella_iori_perello_2009} place fundamentalists, chartists and noise traders in such a system. \cite{farmer_2005} show that noise (Zero-Intelligence) traders submitting orders - limit orders, market orders, and cancellations - at random in a continuous double auction are adequate in explaining 96\% of variance in the bid-ask spread and 76\% of the variance of the price diffusion rate. \cite{cont_stoikov_talreja_2008} propose a stylised model for the dynamics of a limit order book, where order flows are governed by independent poisson processes, separated into independent rates for market orders, limit orders and order cancellations. \cite{gao_2022_hfabm} construct a High Frequency ABM at millisecond levels to simulate the 2010 Flash Crash.

The financial simulation in this paper is based upon these models, with a discrete time LOB operating at a 1-min frequency, populated by Fundamental Traders, Momentum Traders and Noise Traders. Fundamental Traders trade according to an exogenous fundamental value signal, Momentum Traders trade on a momentum signal and Noise Traders place orders at independent rates for market orders, limit orders and order cancellation. The model is calibrated on real market data and includes a calm and a stressed period to investigate resilience and margin procyclicality under stress.

\subsection{Clearing Layer}

Contagion in the model is transmitted through a well documented mechanism. In stressful periods, leveraged institutions need to liquidate in volatile markets to meet margin calls and maintain capital ratios, leading to a procyclical spiral of further calls. \cite{thurner_2012} show that heavy tailed returns and clustered volatility arise from leverage and leverage limits and the liquidation of assets due to these constraints. \cite{aymanns_farmer_2015} further look at the dynamics of these leverage cycles, and find that countercyclical leverage rules dampen the cycle. \cite{bookstaber_2014} make use of an Agent-based model to study the effects of asset and funding based fire sales on financial vulnerability. Together, these papers establish the base of how contagion transmits throught the model, CMs and clients are leverage constrained agents who may need to liquidate during stressful periods. 

\cite{haynes_mcphail_zhu_2019} further state that the Basel III leverage ratio server as a likely binding constraint for derivative positions. \cite{acosta_smith_2018} further show that leverage ratio affected dealers reduce client clearing, restricting clients' access to central clearing. These elements motivate the leverage ratio constraints imposed on Bank and Non-Bank Clearing members and their clients in the Agent-Based model. 

Papers that offer insights into the effect of client clearing and a tiered system are scarce. \cite{paddrik_rajan_young_2020} use CDS trade level data to model how a credit shock triggers margin calls through a network of members and clients, allowing for partial and delayed payments -- a 'soft' default. They find that large client firms in the periphery with imbalanced portfolio contribute most to contagion. \cite{borovkova_2013} find that while CCPs reduce systemic risk in homogenous systems, the effect depends on the system's topology and degree of homogeneity. 


\subsection{Model Calibration}

The credibility of an agent-based model is dependent on how well it is calibrated, i.e. how well it captures realistic behaviours. For the market simulation, the calibration approached used is a minimisation of distance from stylised facts of returns we see on the market. \cite{cont_2001} present a set of stylised facts for returns such as heavy tails, an absence of correlation, volatility clustering and a slow decay of autocorrelation in absolute returns. \cite{cont_2007} find that for agent-based models, volatility clustering in particular remains difficulty to reproduce, and the ones that manage to do so include a switching of the market between high and low activity periods. 

To measure distance from these stylised facts, a loss function must first be defined that captures this robustly. \cite{franke_westerhoff_2012} calibrate a financial market simulation agent-based model by minimising the distance between simulated and empirical moments, by weighing each moment by the inverse of its sampling variability so that more weight is put onto less noisy moments. They construct a sample of these moments using a block-bootrap approach on empirical data. They then validate these moments using moment specific p-values and a moment coverage ratio. Calibration of these moments can be computationally very expensive, as they require a full simulation of the model. \cite{lamperti_2018} address this by pioneer a surrogate modelling approach for the calibration of agent-based models. They train a machine learning surrogate - in this case an XGBoost model - to approximate parameter-to-loss mapping and use active learning to concentrate around the optimum. 

\cite{gao_2022_xgb} further extend this to an XGB-Chiarella approach to estimate the parameters of a Chiarella model. An XGBoost model is trained on a random sample of parameters to predict a loss function. \cite{gao_2022_hfabm} use this methodology to identify a promising region of parameters for their High Frequency Agent-Based market simulation and then perform a grid search around the region to obtain optimal parameters. The thesis makes use of this machine learning surrogate modelling approach to narrow the set of parameters and steer model design. After the model is refined and the set of parameters that need calibration fixed, results from both the XGBoost and Grid Search approach are reported and validated. 