% %==============================================================================
% %  CHAPTER 3 — MODEL STRUCTURE AND DESIGN
% %------------------------------------------------------------------------------
% %  Style: matches introduction.tex conventions (\citep/\citet authoryear; keys
% %  resolve against introduction_refs.bib + model_refs.bib). Calibration of the
% %  behavioural parameters is deliberately deferred to the Data & Calibration
% %  chapter; this chapter fixes the model's structure and cites the design
% %  source for every mechanism (Simudyne ODD first; deviations flagged inline).
% %==============================================================================

This section describes the Client Clearing Agent-Based model and its architecture. The model is built in two tiers - the first is a market simulation where heterogeneous agents trade an asset on a discrete Limit Order Book operating at a one minute frequency, and the second is a clearing layer in which a CCP novates these trades, collects margin calls and handles defaults. The two tiers are connected: price changes and orders in the market layer drive margin calls, trading freezes due to capital ratio requirements, and forced liquidations that impact price and liquidity, creating a loop that spreads contagion. In the spirit of \cite{chiarella_iori_perello_2009} and \cite{majewski_2018}, the market simulation is populated by fundamental traders, momentum traders and noise traders and is calibrated to reproduce stylised facts of returns. Other trader types were trialled to try and make the simulation more realistic. This included splitting the momentum trader into a short-term and long-term momentum trader, adding a Market Maker - both uniform quoting and inventory skewed quoting - and a trader that trades more actively in volatile market situations. These traders, however, did not lead to an improvement in the model and hence added unjustified complexity. 

The data used for calibration is E-mini S\&P 500 (ES) data, but can be calibrated to any other asset. The model includes 2 regimes - a calm regime and a stressed regime, with parameters calibrated separately for each. Parameter values and descriptions can be found in Appendix~\ref{ch:appendix}. In this way, a stressful market environment is produced endogenously. The traders are either Bank Clearing Members trading on their house account, or clients of other Bank and Non-Bank Clearing Members. The clearing layer takes inspiration from the CCP risk model of \cite{deloitte_ccp} and extends it with the client layer, with clearing layer parameters based on regulation and public CCP disclosures. The model is then used to study the effects of client clearing on systemic risk and the procyclicality of margin calls. 


%==============================================================================
\section{Financial Market Simulation}
\label{sec:model-market}

This section presents the design of the financial market simulation:-

\subsection{Limit Order Book}

The market consists of a single asset discrete time Limit Order Book (LOB), with each step corresponding to one minute of trading. At every one minute step, agents submit a limit or market order, with limit orders resting on the book till they are filled, cancelled, or replaced. Orders are matched by price-time priority, and the minimum tick-size corresponds to that of the E-mini: 0.25 index points equating to \$ 12.50 per contract. The mid-price is the average of the best bid and best ask after each auction,
\begin{equation}
  P_t = \frac{b_t + a_t}{2},
\end{equation}
where $b_t$ and $a_t$ are the best bid and best ask at the end of step $t$. The bid-ask spread and liquidity are a product of agent order placement, market environment, and trader balance sheets. 

%------------------------------------------------------------------------------
\subsection{Trading Agents}
\label{sec:agents}

The book is populated by 100 trading agents of three types - 40 fundamental traders (FT), 20 momentum traders (MT), and 40 noise traders/zero-intelligence traders (ZI). The ZIs serve as a constant source of liquidity to the book, the FTs keep asset price anchored to an exogenous fundamental signal, and the MTs amplify price movements and trade on trend. Ten of the FTs also serve as Bank Clearing Members (BCMs), and the remaining 30 and the ZI and MT traders serve as clients of other BCMs and Non-Bank Clearing Members (NBCMs). FTs and MTs place limit orders at each step, with new orders replacing old orders. ZIs place both limit and market orders, and order cancellations, leading to the presence of resting orders on the book. FTs place limit orders at a heterogeneous fundamental signal they observe, and MTs and ZIs place limit orders at a depth of $k$ steps from the previous mid, drawn from a geometric distribution $k \sim \mathrm{Geometric}(p_d)$, where $p_d$ can be fit to L2 order book or calibrated alongside other parameters. A log-normal depth was trialled, but stylised facts were more easily reproduced with a geometric depth of limit order placement.  Order sizes are drawn uniformly from $[1, q_{\max}]$ for each trader.

\subsubsection{Fundamental Trader}

Fundamental traders trade based on the fundamental value of the stock. The fundamental value $V_t$ is fed as an exogenous signal. To introduce heterogeneity, each FT carries a fixed belief offset $z_i \sim \mathcal{N}(0,1)$ drawn at initialisation, so each FT carries its own private valuation. At each step, fundamental traders place a limit order at a reservation $R^{i}_t$ from the fundamental value (FV) signal:- 

\begin{equation}
  R^{i}_t = V_t\,\big(1 + z_i\,c_f\,\sigma_t\big),
\end{equation}
where $\sigma_t$ is the local volatility of the fundamental (Section~\ref{sec:fundamental}) - an EWMA of its squared returns - and
$c_f$ is a calibrated scale that sets the width of the belief
distribution.
The trader buys when its reservation sits above the mid-price and sells when it sits below, placing a limit order at its unique $R^{i}_t$. The heterogeneous $z_i$ being fixed for a trader at onset leads to high $z_i$ traders holding a more optimistic view that causes their positions to accumulate and build, and vice-versa for the low $z_i$ traders. This leads to concentrated inventories, allowing for meaningful exposures to margin calls and market stress. The belief width scale $c_f$ is calibrated as a parameter and controls how far reservations travel from FV and acts as a meaningful lever in reproducing stylised facts, and allows for a more accurate model for the calm and stressed period split. 


\subsubsection{Momentum Trader}

Momentum traders serve as trend followers and track an exponentially weighted moving
average of mid-price log-returns,
\begin{equation}
  M_t = (1-\lambda)\,M_{t-1} + \lambda\,\big(\log P_{t-1} - \log P_{t-2}\big),
\end{equation}
buying when the momentum signal $M_t > 0$ and selling when $M_t < 0$ - placing a limit order on that side at a geometric depth $k \sim \mathrm{Geometric}(p_d)$ from the last mid. Momentum traders amplify trends and contribute to short-horizon volatility clustering. The signal decay $\lambda = 0.05$ is shared across momentum traders and held fixed rather
than calibrated, following \cite{majewski_2018}.
It corresponds to a half-life of about 13.5 minutes, and allowing it to be calibrated did not improve the model. 

\subsubsection{Noise Trader}

Noise, or zero-intelligence (ZI), traders place orders at random and provide the book's
background liquidity \citep{farmer_2005}. Following \cite{cont_stoikov_talreja_2008}, ZI order flow is modelled as unique rates for limit order, market order and order cancellation placements, following a Poisson process, corresponding to a per-step probability. 
 A ZI trader therefore submits a limit order with probability $\alpha_{zi}$ on a random side at a geometric
depth $k \sim \mathrm{Geometric}(p_d)$, submits a market order with probability $\mu_{zi}$ on a
random side, and cancels each resting order with probability $\delta_{zi}$. Unlike FT and MT, ZI orders stay on the book until they are filled or cancelled. The three rates $\alpha_{zi}$,
$\mu_{zi}$, and $\delta_{zi}$ are calibrated as part of the surrogate modelling loop.

\subsection{The Fundamental Value Process}
\label{sec:fundamental}

The fundamental value $V_t$ is exogenously created and fed as a signal to fundamental traders (FTs). By default, the fundamental value is taken as the empirical mid price series for E-mini (ES) futures. A synthetic GBM with jumps and a mean reverting volatility process is kept as an alternative, and both lead to a realistic reproduction of stylised facts.


\paragraph{Data-driven signal.}
By default $V_t$ is the observed one-minute ES mid, and overnight gaps are maintained as they drive a large portion of returns and contribute to stress. This also allows for the FV signal to directly replicate the properties of the calm and stressed period.

\paragraph{Synthetic alternative.}
The alternative approach replaces the real series with a synthetic series, a Geometric Brownian
Motion whose instantaneous volatility $\sigma_t$ follows a mean-reverting Ornstein--Uhlenbeck
process \citep{stein_stein_1991,heston_1993} overlaid with Merton jumps \citep{merton_1976}:
\begin{align}
  \sigma_{t+1}            &= \theta + e^{-\kappa}\,(\sigma_t - \theta) + \xi\,\varepsilon^{\sigma}_t, \\
  \log V_{t+1} - \log V_t &= \mu + \sigma_t\,\varepsilon_t + J_t,
      \qquad J_t \sim \mathcal{N}(0,\sigma_J^2).
\end{align}
The jumps $J_t$ arrive as a Poisson process---roughly two per session in the calm regime and one
in the stressed regime---with sizes drawn from a normal distribution whose standard deviation
$\sigma_J$ is small, about $0.1\%$ per jump in calm markets and $0.7\%$ under stress. The
mean-reverting volatility is the feature that generates realistic long-horizon volatility
clustering in the model.
\paragraph{Overnight gaps.}
Both fundamentals retain overnight gaps, which holds relevance as the largest moves in the stressed sample occurred overnight. The data-driven path carries the realised gaps
in sequence, while the synthetic path draws each gap from the regime's observed overnight returns.

\paragraph{Local volatility.}
Each fundamental is accompanied by a local volatility $\sigma_t$ that feeds into the FT reservation. For the synthetic path this is simply the mean-reverting volatility above. For
the data-driven path it is an exponentially weighted moving average of $V_t$'s squared one-minute
returns,
\begin{equation}
  \sigma_t^2 = (1-\phi)\,\sigma_{t-1}^2 + \phi\,\big(\log V_t - \log V_{t-1}\big)^2,
\end{equation}
with the weight $\phi$ averaging over a short intraday window and the average reset at each session open so the overnight gap does not enter it. 





% %==============================================================================
% \section{The margin framework}
% \label{sec:model-margin}

% The clearing tier connects the market to a single CCP through the star
% topology of \citet{simudyne_odd}: every clearing member holds a bilateral
% link to the CCP, which becomes the counterparty to all cleared positions
% upon novation. Two member types are distinguished, as in the ODD. A
% \emph{banking clearing member} (BCM) trades its own account --- with
% fundamental-trader logic, so the market-layer calibration is unaffected by
% the cast --- and additionally guarantees a book of clients. A
% \emph{non-banking clearing member} (NBCM) is a pure intermediary with no
% proprietary position; its entire balance-sheet capacity is client clearing.
% The thesis extension is the client tier itself: ninety end-clients
% (fundamental, momentum, and zero-intelligence types) clear through the ten
% client-carrying members in skewed books of five to fifteen clients, with the
% non-banking members carrying the largest books. Loss-absorbing capital is
% initialised at regulatory scale --- bank-level capital for BCMs, the
% non-bank futures-commission-merchant range for NBCMs, and entity-type-scaled
% cash for clients --- taken from CFTC member financial data and CCP
% disclosures.

% \paragraph{Variation margin.}
% Every 60 minutes the margin cycle marks each cleared book to the prevailing
% mid and settles the move in cash. Settlement is computed against
% \emph{average filled prices}, the ODD's margin-call profit-and-loss
% convention \citep{simudyne_odd}: writing $N_\tau$ for the end-of-cycle
% position, $m_\tau$ for the mark, and $\{(q_k, p_k)\}$ for the fills executed
% during the cycle,
% \begin{equation}
% \label{eq:vm}
% \mathrm{VM}_\tau \;=\;
% \Bigl[\, N_\tau\,(m_\tau - m_{\tau-1})
% \;+\; \textstyle\sum_k q_k\,(m_{\tau-1} - p_k) \Bigr]\cdot \ell \cdot \chi ,
% \end{equation}
% where $\ell$ is the per-regime contract lot that maps model volume to
% empirical ES volume and $\chi = \$50$ per index point is the CME multiplier.
% The second term in \eqref{eq:vm} settles each trade from its execution
% price, so that the slippage of forced liquidations
% (Section~\ref{sec:model-default}) is borne in cash by the liquidating party
% rather than lost between marks. Cleared participants use futures-style
% accounting throughout: a fill moves only the position, and cash moves only
% through the margin cycle.

% \paragraph{Initial margin.}
% Initial margin is charged on gross exposure as a value-at-risk scan --- for
% a single linear future, the SPAN scanning charge coincides with parametric
% VaR --- per EMIR's requirements for exchange-traded derivatives
% \citep{emir_rts}:
% \begin{equation}
% \label{eq:im}
% f^{\mathrm{IM}}_\tau \;=\;
% \max\!\Bigl( f_{\mathrm{floor}},\; z_{99}\,\hat\sigma^{\mathrm{day}}_\tau
% \sqrt{h} \Bigr),
% \qquad
% \mathrm{IM}_\tau = f^{\mathrm{IM}}_\tau \cdot
% \bigl(\,|N^{\mathrm{own}}_\tau| + \textstyle\sum_{c}|N^{c}_\tau|\,\bigr)
% \cdot \ell\,\chi\, m_\tau ,
% \end{equation}
% with $z_{99}$ the 99\% normal quantile, $h = 2$ days the margin period of
% risk, and an anti-procyclicality floor $f_{\mathrm{floor}}$ anchored to the
% CME's ES margin scale. The driving volatility $\hat\sigma^{\mathrm{day}}_\tau$
% defines the \emph{margin regime}, the experimental lever for the
% procyclicality hypothesis: in the \emph{reactive} baseline it is an
% exponentially weighted estimate of realised intraday simulation returns
% (half-life two sessions), so margin ratchets upward through a crash exactly
% as CME's ES margin was raised repeatedly through March 2020; a \emph{static}
% variant fixes it at the regime constant; and a \emph{flat} variant removes
% the volatility response altogether, charging a constant fraction --- the
% through-the-cycle comparator contemplated by the anti-procyclicality
% options of EMIR and the post-2020 margin debate
% \citep{murphy_2014,esrb_2020}. A maintenance threshold at $95\%$ of initial
% margin flags margin calls when a cycle's variation margin exceeds the
% remaining buffer \citep{simudyne_odd}.

% \paragraph{Capital adequacy.}
% Each member's capital ratio is cash over initial margin,
% \begin{equation}
% \label{eq:capital}
% \kappa_\tau \;=\; \frac{C_\tau}{\mathrm{IM}_\tau} \;\geq\; 8\%,
% \end{equation}
% the futures-commission-merchant net-capital rule, which sets required
% capital against \emph{risk margin} rather than gross notional
% \citep{cftc_reg117}. The distinction matters operationally: an FCM clears
% many multiples of its capital in client notional, so a notional-based floor
% would bind permanently, whereas \eqref{eq:capital} binds only as distress
% approaches. A banking member's proprietary book is additionally capped by a
% trading-book VaR limit, $|N^{\mathrm{own}}|\,\ell\chi m \leq
% \beta C / (z_{99}\hat\sigma^{\mathrm{day}})$ with $\beta = 5\%$ of capital,
% in the spirit of regulatory trading-book risk limits \citep{basel_frtb};
% without it, the fundamental-trader logic accumulates unbounded prop
% positions in a trend.

% \paragraph{Client clearing.}
% Each client carries its own balance sheet. The margin cycle issues every
% client its own variation-margin call, paid from the client's own cash and
% relayed by its member to the CCP; the client's live inventory is novated
% into the member's cleared book each cycle. A client can open positions only
% up to its broker house margin --- a fixed $20\%$ of notional, i.e.\
% five-times leverage, the conservative house multiple above the exchange
% minimum --- and is frozen (barred from new risk) when its cash falls below
% $8\%$ of its gross notional. Cash exhaustion is a client default and hands
% the position to the member (Section~\ref{sec:model-default}). This
% client tier is the thesis's central extension of the reference model;
% its empirical motivation --- clients supply the majority of margin at major
% CCPs and typically depend on a single clearing agent --- is documented by
% \citet{ofr_2026}.

% %==============================================================================
% \section{The default management framework}
% \label{sec:model-default}

% Default management proceeds in three escalating stages --- client default,
% member distress, member default --- and is closed by the CCP's prefunded
% resources. All liquidations use the same execution schedule: a position of
% size $X$ unwound over horizon $H$ follows the closed-form
% \citet{almgren_chriss_2000} trajectory
% \begin{equation}
% \label{eq:ac}
% \frac{x_j}{X} \;=\; \frac{\sinh\bigl(\kappa (H - j)\bigr)}{\sinh(\kappa H)},
% \qquad j = 0,\dots,H,
% \end{equation}
% with urgency $\kappa H = 2$ and $H = 30$ minutes, executed as market orders
% on the simulated book --- so liquidation impact is endogenous: the fire-sale
% walks the very book on which surviving members are marked, which is the
% price-mediated contagion channel of the model
% \citep{simudyne_odd,vytelingum_2025}.

% \paragraph{Client default.}
% A client whose cash is exhausted by variation margin defaults. Its member
% covers the uncollected margin call, assumes the client's position onto its
% own book, and liquidates it by \eqref{eq:ac} --- a banking member through
% its own orders, a non-banking member through orders routed by the CCP on its
% behalf. Because variation margin settles against fill prices
% \eqref{eq:vm}, the close-out loss is realised as the fire-sale walks the
% price: the member's loss is the deficit the liquidation actually produces,
% not a stylised haircut. This client$\to$member loss assumption is the first
% link of the cascade under study.

% \paragraph{Member distress.}
% A member breaching the capital floor \eqref{eq:capital} while still solvent
% acts before defaulting, asymmetrically by type as in the ODD
% \citep{simudyne_odd}: a banking member \emph{deleverages}, selling its
% proprietary book down by \eqref{eq:ac} until the floor is restored; a
% non-banking member, having no own book to sell, \emph{stops out} ---
% suspending new client risk, which freezes its clients --- for as long as the
% breach persists. The stop-out is state-based: it lifts if the ratio
% recovers. These are the liquidity-withdrawal behaviours that couple margin
% pressure back into the market layer.

% \paragraph{Member default and the waterfall.}
% A member whose cash is exhausted defaults. The CCP absorbs the member's
% realised cash deficit --- the shortfall it owes but cannot pay --- through
% the five-level waterfall of the reference model \citep{simudyne_odd}, in
% EMIR's defaulter-pays ordering \citep{emir_rts}:
% \begin{equation}
% \label{eq:waterfall}
% \text{L1 own DF contribution}
% \;\to\; \text{L2 CCP skin-in-the-game}
% \;\to\; \text{L3 pooled DF}
% \;\to\; \text{L4 survivors' cash}
% \;\to\; \text{L5 CCP capital}.
% \end{equation}
% The defaulted member is then removed from the system (ODD step sequence).
% Its surviving clients are \emph{ported}: each is reassigned to a surviving
% client-carrying member with spare risk capacity, accepted only if the
% receiver remains above its capital floor with the transferred book ---
% porting is the regulatory first option on a member default
% \citep{emir_rts}, but it is capacity-checked rather than guaranteed, since
% in practice the modal client has no standing backup agent
% \citep{ofr_2026}; clients that cannot be placed are closed out and
% deactivated. The member's own book, together with any unported client
% positions, passes to the CCP, which disposes of it by \eqref{eq:ac} on the
% open market. The CCP carries this book in a disposal account marked through
% the margin cycle: disposal-period losses re-enter the waterfall
% \eqref{eq:waterfall} of the originating default, so the mutualised loss is
% the close-out loss actually realised, symmetric with the client treatment.
% The open-market disposal is a flagged deviation from the ODD, whose default
% positions are auctioned to the surviving member with the largest opposing
% position; the deviation is deliberate, as the endogenous price impact of
% disposal is precisely the contagion mechanism the thesis measures.

% \paragraph{The default fund.}
% The mutualised layers of \eqref{eq:waterfall} are prefunded. The fund is
% sized daily to the \emph{cover-two} standard --- sufficient for the
% simultaneous default of the two largest members
% \citep{emir_rts} --- by the stress-loss-over-initial-margin method of
% Euronext Clearing \citep{euronext_a9}:
% \begin{equation}
% \label{eq:df}
% \mathrm{DF}_\tau \;=\; (1+b) \sum_{j \in \text{top-2}}
% \max\Bigl( 0,\; \bigl(s_\tau - f^{\mathrm{IM}}_\tau\bigr)\,
% |N_{j,\tau}|\,\ell\,\chi\,m_\tau \Bigr),
% \qquad
% s_\tau = \max\bigl(15\%,\; z_{s}\,\hat\sigma^{\mathrm{day}}_\tau\sqrt{h}\bigr),
% \end{equation}
% the loss each member would suffer beyond posted margin under an
% extreme-but-plausible move $s_\tau$, with buffer $b = 10\%$. The CCP
% prefunds a skin-in-the-game share ahead of the pooled fund, consumed at L2
% before any survivor resources \citep{simudyne_odd}; the EMIR skin-in-the-game
% rule is expressed against CCP regulatory capital rather than fund size, so
% the model's fund-share convention is a stylised assumption and is disclosed
% as such. Member contributions are paid \emph{in cash} into the fund, pro
% rata to exposure, and every post-draw daily recomputation is therefore a
% replenishment cash call on survivors --- the margin-liquidity channel that
% made the March 2020 episode a funding event as much as a solvency one
% \citep{esrb_2020}. Reverse-stress experiments scale the fundamental path
% until \eqref{eq:df} is breached, the scenario-design logic of
% \citet{euronext_a9}.

% Two structural counterfactuals complete the design and operationalise the
% research questions: a \emph{direct-clearing} configuration, in which the
% client tier is removed and all end-users face the CCP directly with
% otherwise identical trading mechanics --- isolating the loss-absorption
% structure of tiering in the spirit of the netting and tiering trade-offs of
% \citet{duffie_zhu_2011} and \citet{galbiati_soramaki_2013} --- and the
% margin-regime switch of Section~\ref{sec:model-margin}. Both are evaluated
% as paired experiments on identical fundamental paths and random seeds, so
% that differences in outcomes are attributable to the structural change
% alone.