% %==============================================================================
% %  CHAPTER 4 — DATA, CALIBRATION AND VALIDATION (methodology)
% %------------------------------------------------------------------------------
% %  Style: matches introduction.tex / model.tex (\citep/\citet authoryear; keys
% %  resolve against introduction_refs.bib + model_refs.bib). Written while the
% %  final calibration runs are now LOCKED (D65, output/relock/): this chapter fixes the
% %  METHODOLOGY; calibrated values, fit tables, loss surfaces, and the ablation/robustness
% %  numbers are reported in the results chapter. [NUMBERS: ...] marks the spots
% %  to fill from output/relock/ (and output/relock_gbm/ for the SV-MJD robustness).
% %==============================================================================

This section presents the methodology for calibrating the agent based financial market simulation. As this serves as the first stage and base of the clearing layer, the market simulation should be well calibrated and able to reproduce the features of real market returns. A common method to do this, and the one adopted by this model, is the reproduction of the moments of empirical returns through the simulation. This chapter describes the data used, the parameters calibrated, the method of calibration and the calibration, and a validation of the calibrated parameters. The chosen moments are the stylised facts of returns we see empirically - the distribution of returns, heavy tails, absence of autocorrelation and volatility clustering. 

\section{Data and Regimes}

The dataset used for calibration is the E-mini S\&P futures at a one minute frequency, matching the model time steps. There are two separate regimes, a calm regime and a stressed regime. The calm regime spans the a calm stretch in late 2019, and the stressed regime includes data from the early Covid-19 window, which was referred to as the most significant test of the resilience of financial markets since the GFC \citep{bcbs_cpmi_iosco_2022}. Parameters are calibrated separately on both regimes, as they differ substantially in volatility, microstructure and return distribution. Returns are intraday log-returns of the mid price, and some summary statistics are reported in \ref{tab:cal-data}.

\begin{table}[h] 
\centering
\small
\setlength{\tabcolsep}{5pt}
\caption{\textbf{Empirical summary statistics for the two calibration regimes.}}
\label{tab:cal-data}
\begin{threeparttable}
\begin{tabular}{lccccccccc}
\toprule
 & & \multicolumn{4}{c}{One-minute returns} & \multicolumn{4}{c}{Daily returns}\\
\cmidrule(lr){3-6}\cmidrule(lr){7-10}
Regime & Sessions & Mean & SD & Skew. & Ex.\ kurt. & Mean & SD & Skew. & Ex.\ kurt.\\
 & & (bps) & (bps) & & & (\%) & (\%) & &\\
\midrule
Calm (2019) & 75 & 0.00 & 2.40 & $-1.41$ & 55.0 & 0.10 & 0.60 & $-0.85$ & 1.66 \\
Stressed (2020) & 73 & $-0.03$ & 12.40 & $-1.24$ & 67.1 & $-0.13$ & 3.14 & $-0.31$ & 2.50 \\
\bottomrule
\end{tabular}
\begin{tablenotes}[flushleft]\footnotesize
\item One-minute returns exclude overnight gaps; daily returns are close-to-close. bps$=10^{-4}$.
\item Both regimes hold $\approx$29{,}000 one-minute observations (29{,}961 vs 28{,}866) over a
comparable number of full RTH sessions (75 vs 73), so the two samples are of matched size.
\end{tablenotes}
\end{threeparttable}
\end{table}

Limit order depth placement follows a geometric law calibrated besides the other parameters, but this can also be directly estimated from L2 data. 

\section{Loss Function: Stylised Facts Distance}
\label{sec:cal-loss}

The market layer is calibrated to reproduce the stylised facts of returns we see in markets \citep{cont_2001} such as heavy tails, near-zero autocorrelation, and volatility clustering. Five different moments are used as proxies for these stylised facts, and are computed on simulated and empirical one minute log-returns. Parameters are chosen to reduce the distance between a weighted loss function defining these stylised facts in the simulation and in empirical data, similar to \cite{gao_2022_xgb}. The function there is further extended by the Hill Tail index used in \cite{gao_2022_hfabm}, and the weighted loss function takes the form:
\begin{equation}
\label{eq:cal-D}
D(\theta) = w_{\mathrm{KS}}\,\mathrm{KS}(\theta) + w_{V}\,\Delta V(\theta)
          + w_{\mathrm{ACF1}}\,\Delta\mathrm{ACF1}(\theta)
          + w_{\mathrm{ACF2}}\,\Delta\mathrm{ACF2}(\theta)
          + w_{\mathrm{Hill}}\,\Delta\mathrm{Hill}(\theta).
\end{equation}
The calibrated parameter vector is the minimisation of this distance over the admissible set $\Theta$,
\begin{equation}
\label{eq:cal-argmin}
\hat\theta = \arg\min_{\theta\in
\Theta} D(\theta).
\end{equation} 

\noindent A description of the five stylised facts are given below.

\paragraph{Distribution (KS).} The Kolmogorov--Smirnov statistic, from the KS test which tests for the equality of probability distributions.
$\mathrm{KS}=\sup_x|F_{\mathrm{sim}}(x)-F_{\mathrm{emp}}(x)|$ is the largest gap (supremum) between the simulated
and empirical return distribution functions. Matching it forces the simulated returns to share the
distribution function of the empirical distribution, which is an estimate of the cumulative distribution function \citep{gao_2022_xgb}.

\paragraph{Volatility ($\Delta V$).} The absolute difference in return standard deviation between the empirical and simulated returns
\begin{equation}
\Delta V = \big| \sigma_{\mathrm{sim}} - \sigma_{\mathrm{emp}} \big|,
\end{equation}
where $\sigma$ denotes the standard deviation of the one-minute return series.

\paragraph{Return autocorrelation ($\Delta\mathrm{ACF1}$).} Empirical returns have near-zero autocorrelation. To reproduce this, we take the mean absolute difference between the simulated and empirical return autocorrelations over the lags $L=\{1,5,10,20\}$ minutes,
\begin{equation}
\Delta\mathrm{ACF1} = \frac{\sum_{\ell\in L}\big|\,\mathrm{ACF}_{sim}(\ell,r)-\mathrm{ACF}_{emp}(\ell,r)\,\big|}{|L|}.
\end{equation}
Following \citet{gao_2022_xgb}, each $\mathrm{ACF}(\ell,r)$ is smoothed over three adjacent lags,
$\mathrm{ACF}(\ell,r)=\tfrac13(a_\ell+a_{\ell+1}+a_{\ell+2})$ with $a_k$ the lag-$k$ return
autocorrelation, so lags $\{1,2,3,\,5,6,7,\,10,11,12,\,20,21,22\}$ enter the function. The
component penalises predictability such as a bid-ask bounce.


\paragraph{Volatility clustering ($\Delta\mathrm{ACF2}$).} In empirical data, we see that volatility clusters over time, large returns are followed by large returns and low returns by low returns. This can be measure by studying the autocorrelation of absolute or squared returns. Autocorrelations of absolute returns are positive and decay slowly. $\Delta\mathrm{ACF2}$ is the same distance with the autocorrelations $\mathrm{ACF}(\ell,|r|)$
taken on $|r|$ (same lags, same three-lag smoothing),
\begin{equation}
\Delta\mathrm{ACF2} = \frac{\sum_{\ell\in L}\big|\,\mathrm{ACF}_{sim}(\ell,|r|)-\mathrm{ACF}_{emp}(\ell,|r|)\,\big|}{|L|}.
\end{equation}
For the dataset, squared returns were dominated by extreme values making their autocorrelation noisy, and hence absolute returns are used

\paragraph{Heavy Tails (Hill).} Tail heaviness is measured by the \citet{hill_1975} estimator of the
tail index of absolute returns, $\mathrm{Hill}$ (a lower value means a fatter tail), and its absolute difference is taken,
\begin{equation}
\Delta\mathrm{Hill} = \big| \mathrm{Hill}_{\mathrm{sim}} - \mathrm{Hill}_{\mathrm{emp}} \big|.
\end{equation}
The estimate is averaged over several tail cut-offs for robustness. \\

This combination of the KS statistic and Hill tail index is deliberate, using just the KS statistic makes tails very large and using just the Hill Tail index makes kurtosis unrealistically high. This is a deviation from \cite{gao_2022_hfabm} that use only the Hill tail index for their high-frequency LOB and \cite{gao_2022_xgb} that use only the KS statistic, and is motivated by the more granular step size alongside the importance of realistic tails due to the study of stressed periods.

\paragraph{Weighting.} The five loss components are each of varying magnitudes. To make them of comparable scale, such that the largest scale term does not dominate, the components are weighed by the reciprocal of their sampling variability following the work of \cite{franke_westerhoff_2012}. Each weight is the reciprocal of the component's standard deviation,
\begin{equation}
\label{eq:cal-w}
w_c = \frac{1}{s_c}, \qquad c \in \{\mathrm{KS},\,V,\,\mathrm{ACF1},\,\mathrm{ACF2},\,\mathrm{Hill}\},
\end{equation}
such that each term is measured in the units of it's sample noise. This has the added benefit that down-weighting noisy moments such that the loss function offers higher weight to moments that are more consistently identified across the sample. 
The $s_c$ come from a moving-block bootstrap of the empirical returns \citep{kunsch_1989}, day long blocks are taken as independent samples, and the moments are recomputed on each resample/block. For the multi-lag components $\mathrm{ACF1}$ and $\mathrm{ACF2}$ the reciprocal-SD
standardisation is applied to each lag inside the average. \cite{franke_westerhoff_2012} made use of the full inverse-covariance matrix for weighting and \cite{gao_2022_hfabm} made use of inverse variance, as the former is near singular for these losses and for the latter the return standard deviation dominates due to low sampling variability, and hence the inverse of the standard deviation is employed. 


\section{Calibration Methodology}
\label{sec:cal-method}

The ABM consists of several parameters that need calibration. Running a multi-day simulation over several random parameter sets to find an optimal set which minimises the loss function would be computationally very expensive. For this model, seven parameters need calibration across two different regimes. The paper therefore uses a surrogate modelling technique to lower this computational cost. The technique was  pioneered by \cite{lamperti_2018}, who train an XGBoost model to classify positive and negative calibrations to find an optimal set of parameters for an Agent Based Model. It was extended by \cite{gao_2022_xgb}, who train the XGBoost model to predict the loss function for a given set of parameter in place of running the simulator.

\subsection{Surrogate Modelling}

The surrogate model used for calibration is an XGBoost model, that is trained on the mapping from $\theta\mapsto D(\theta)$, from parameter vector to loss. Calibration follows the two stage approach of \cite{gao_2022_hfabm}, where the first stage is an estimation of an optimal set of parameters using the surrogate model approach of \cite{gao_2022_xgb}, and the second stage a grid search around those optimal parameters in a feasible bound. The model parameter values that minimise the stylised facts distance are then chosen. The following steps lay out the calibration methodology:
\begin{enumerate}
\item \textbf{Sampling.} A large pool of parameter combinations is drawn using Sobol Sampling, to ensure a uniform distribution across the multi-dimensional space of parameters. A random set of these parameters is then taken and the ABM simulation run on them to get a stylised facts distance as a label. This then gives us an initial training set of input parameter combinations and their corresponding stylised facts distances as labels.
\item \textbf{Model Training.} An XGBoost regressor is trained on these $(\theta, D)$ pairs, to take parameter combinations as an input and predict stylised facts distance as an output.
\item \textbf{Search and supplement.} The cheap emulator predicts the loss over a large unevaluated
Sobol pool; the most promising vectors are then run on the true simulator and added to the training
set. The selection of the training set combines a selection with the smallest predicted loss (exploitation), and a randomly drawn fraction (exploration) to avoid possible bias. 
\item \textbf{Iterate (stage one).} A new XGBoost model is trained on this training set, and this cycle is repeated countless times until it returns the surrogate-optimal parameter vector.
\item \textbf{Refine (stage two).} A tight grid search is then performed around a small range centred on the surrogate optimum, and the true simulator is used to evaluate this. The node with the lowest stylised facts distance is then taken as the calibrated set of parameters  $\theta^\star$. The located $\theta^\star$ is then validated on fresh seeds (\S\ref{sec:cal-validation}).
\end{enumerate}

The quality of the surrogate model is reported using its $R^2$, with a high $R^2$
indicating that the surrogate emulated the model with high accuracy. The model had an $R^2$ of 0.90 for the calm regime and !!!! for the stressed regime. 


\begin{table}[h]
      \centering
      \caption{Model parameters: status and values. The seven trading-agent behavioural parameters are
      calibrated per regime; the optima (marked ---) are filled from the surrogate re-lock. The primary
      specification takes $V_t$ to be the contemporaneous empirical one-minute mid (no free parameters);
      the SV-MJD parameters below are measured from realized measures and govern the robustness arm only.
      Structural quantities are fixed by design or contract.}
      \label{tab:model-params}
      \begin{tabular}{lllcc}
      \hline
      Parameter & Description & Status & Calm & Stressed \\
      \hline
      \multicolumn{5}{l}{\emph{Trading agents -- calibrated (seven per regime)}} \\
      $c_f$          & FT belief-width scale       & Calibrated & --- & --- \\
      $\alpha_{zi}$  & ZI limit-order rate         & Calibrated & --- & --- \\
      $\mu_{zi}$     & ZI market-order rate        & Calibrated & --- & --- \\
      $\delta_{zi}$  & ZI cancellation rate        & Calibrated & --- & --- \\
      $p_d$          & geometric placement depth   & Calibrated & --- & --- \\
      $\lambda$      & MT signal (EWMA) decay      & Calibrated & --- & --- \\
      $\gamma_M$     & MT activation scale         & Calibrated & --- & --- \\
      \hline
      \multicolumn{5}{l}{\emph{Fundamental -- SV-MJD path (realized measures)}} \\
      $\sigma_d$     & diffusion                   & Measured & $2.88\times10^{-4}$ & $1.80\times10^{-3}$ \\
      $\lambda_J$    & jump intensity (per day)    & Measured & 1.93    & 0.97    \\
      $\delta$       & jump-size std               & Measured & $1.12\times10^{-3}$ & $7.26\times10^{-3}$ \\
      $\alpha$       & OU mean-reversion           & Measured & $5.7\times10^{-3}$  & $2.7\times10^{-3}$  \\
      $\theta$       & OU long-run volatility      & Measured & $2.34\times10^{-4}$ & $1.48\times10^{-3}$ \\
      $s$            & OU volatility-of-volatility & Measured & $1.49\times10^{-5}$ & $4.88\times10^{-5}$ \\
      $\mu$          & drift                       & Measured & $1\times10^{-6}$ & $-6\times10^{-6}$ \\
      \hline
      \end{tabular}
      \end{table}



\section{Validation}
\label{sec:cal-validation}

This section presents the results of the financial market simulation, and the efficacy of the calibration. The Moment Coverage Ratio (MCR) of \cite{franke_westerhoff_2012} is used as a check, which checks how well each particular moment is matched, unlike the loss function which aggregates them. The MCR creates a 95\% confidence interval for each moment with boundaries ±1.96 times standard deviation around empirical value. The standard deviation is obtained by once again applying the block-bootstrap method on empirical data and getting a moment value for each block/sample, and taking it's standard deviation. The MCR is then the proportion of simulation runs for which the moment is contained in the confidence interval, an approach employed by \cite{gao_2022_hfabm}.

\begin{table}[h]
\centering
\small
\caption{\textbf{Moment coverage ratio (MCR) by moment.} Fraction of independent simulation runs whose
moment falls inside the 95\% block-bootstrap confidence interval of the empirical moment.}
\label{tab:cal-mcr}
\begin{threeparttable}
\begin{tabular}{lcc}
\toprule
Moment & Calm & Stressed \\
\midrule
Return SD            & 100\% & --- \\
Hill tail index      & 92\%  & --- \\
ACF $r(1)$           & 46\%  & --- \\
ACF $r(5)$           & 100\% & --- \\
ACF $r(10)$          & 100\% & --- \\
ACF $r(20)$          & 100\% & --- \\
ACF $|r|(1)$         & 98\%  & --- \\
ACF $|r|(5)$         & 0\%   & --- \\
ACF $|r|(10)$        & 0\%   & --- \\
ACF $|r|(20)$        & 0\%   & --- \\
\midrule
Mean per-moment      & 63.6\% & --- \\
\bottomrule
\end{tabular}
\begin{tablenotes}[flushleft]\footnotesize
\item The fraction of independent runs whose moment lands in the 95\% empirical interval.
\end{tablenotes}
\end{threeparttable}
\end{table}

For the calm regime, as can be seen, some moments are very accurately estimated. Long-horizon volatility clustering, however, appears to be structurally unreachable. We see weaker volatility clustering in the model than we see in empirical returns, and this is more prominent for the calm regime. For the stressed regime,  The Kolmogorov-Smirnov statistic is not included in the Moment Coverage Ratio as it has no empirical value or confidence interval to fall under. The distributional fit can be seen through the CDF of returns.


